---
title: "What are the effects of education expenditure on Urban and Rural Populations?"
author: "Luit Deka"
execute: 
  echo: false
---

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(rstanarm)
library(primer.data)
```

```{r}
#| message: false
#| warning: false
dataset <- read_csv("Data.csv") |>
  as_tibble() |>
  filter(Time > 2009)
```

```{r}
#| message: false
#| warning: false
x <- dataset |>
  mutate(Rural_PopPerc = `Rural population (% of total population) [SP.RUR.TOTL.ZS]`, Urban_PopPerc = `Urban population (% of total population) [SP.URB.TOTL.IN.ZS]`, Rural_PopGrowth = `Rural population growth (annual %) [SP.RUR.TOTL.ZG]`, Urban_PopGrowth = `Urban population growth (annual %) [SP.URB.GROW]`, Educ_Exp_Perc = `Government expenditure on education, total (% of government expenditure) [SE.XPD.TOTL.GB.ZS]`, Country_Name = `Country Name`)|>
  select(Time, Country_Name,
         Rural_PopPerc, Rural_PopGrowth, 
         Urban_PopPerc, Urban_PopGrowth,
         Educ_Exp_Perc)

```

How does the effects of national government expenditure on education differ from the urban population to the rural population? The data to answer the question was sourced from the World Bank Universal Indicator dataset, and then the relevant data frames were combined to analyze the relationship between government expenditure on urban populations vs. rural populations. Increased expenditure on education typically increases the urban population, decreases the rural population, but still benefits both in terms of quality of life and access to resources.






